\subsection{Tetris configuration}

When finding the optimal settings for the Cross-entropy method (section \ref{optimalsettingsce})
and CMA-ES (section \ref{optimalsettingscma}), we used a game complexity and feature set 
setting which resulted in the fastest runtime. We had to run over 6,000 individual experiments
in the process of finding the optimal settings, so with our limited time frame, we had
to use fastest settings.
If we had a longer time frame, we would preferable have used normal Tetris instead
of hard Tetris when finding the optimal settings. Even though our game complexity experiments
(section \ref{HardTetris}) indicated that the difficulty of the game doesn't signify a difference, it would have been more correct to use normal Tetris for finding the optimal settings, since
we use normal Tetris in the final comparison experiments. Though, it could possible be that the
optimal setting for hard Tetris is the same optimal setting for normal Tetris.
The same thing can be said about the feature sets. In the majority of our experiments, we
use the Bertsekas feature set, even though that the Dellacherie feature set results in
higher scores. But as our experiments from section \ref{compoffeatureset} revealed, the
scores only seems to be shifting and therefore not creating a performance difference for
the Cross-entropy method or CMA-ES algorithm. However, we would have preferred that we could have
executed our final comparison experiments with the Dellacherie feature set, namely because
it's possible to achieve higher scores with this feature set, compared to the Bertsekas feature
set.

