\documentclass[11pt,a4paper]{article}
\usepackage{a4wide}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{amsfonts}

\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[tikz]{bclogo}
\usepackage[framemethod=tikz]{mdframed}

\mdfdefinestyle{mystyle}{%
  rightline=true,
  innerleftmargin=10,
  innerrightmargin=10,
  outerlinewidth=3pt,
  topline=false,
  rightline=true,
  bottomline=false,
  skipabove=\topsep,
  skipbelow=\topsep
}

\setlength{\parindent}{0cm}

\begin{document}

\title{BSc Project Description\\\textbf{Learning to Play Tetris using
    the Covariance Matrix Adaptation
    Evolution Strategy}}
\date{\today}
\maketitle

% Removed because Motivation and background is the same
%\input{Background}

\input{Motivation}

\input{Scope}

\input{PrevWork}

\input{Approach}

\input{Requirements}

\input{Supervision}

\input{LearningGoals}

\bibliography{BSc2014}[]
\bibliographystyle{plain}


\end{document}

\title{BSc Project Description\\\textbf{Learning to Play Tetris using
    the Covariance Matrix Adaptation
    Evolution Strategy}}
\date{\today}
\maketitle

\section{Time plan}

\subsection{Unofficial - pre-schedule}
This is the time before the official start 
of our bachelor project, i.e. now.\\\\
Setup Shark: Done, on UNIX systems.\\\\
Setup MDPTetris: before 19-10-2015\\\\
Sample CMA-ES simulations,
to get a feel of complexity and runtime: before 16-11-2015

\subsection{Official bachelor schedule}
Upload contract: 20-11-2015\\\\
Uplaod Synpsis: 04-12-2015\\\\
Implement CE for Shark/Tetris: before 10-12-2015\\\\
Setup both experiments: before 15-12-2015\\\\
Upload report: 18-01-2016\\\\
Defense: (28 | 29)-01-2016


\section{Problem statement}
In comparison between the Cross-entropy Method and
Covariance Matrix Adaption Evolution Strategy, which of the
two performs better in training an agent for playing Tetris?

\section{Learning goals}

\subsection{Reinforcement learning}
Investigate how to apply reinforcementlearning 
to the task of playing Tetris.

\subsection{Neural network}
Learn how to setup and apply Neural Networks as an agent for playing
Tetris.

\subsection{CMA-ES and CE}
Learn the fundamentals of genetic evolutionary algorithms
and in particular how CMA-ES end CE works.

\section{Goals}

\subsection{Descriibe and implement Cross-Entropy}
Obtain an implementation of the CEM (Cross-Entropy Method) either 
by implementation or by usage of third-party source. 

\subsection{Implement interface between Shark and Tetris}
Find a suitable tetris playing agent which can be optimized 
by using either CMA-ES or CEM. Implement interface between 
the trainable agent and a Tetris simulator.

\subsection{Compare the two methods}
Run each of the optimization methods and compare
the results to determine weather one outperforms the 
other.

\section{Limitations}

\subsection{Running enviroment}
The implementation will be tested and run 
on a UNIX system, namely Ubuntu 14, however, 
the implementation should be able to run on 
any Unix distribution with the applied setup as in the project. 

\subsection{Simulation configurations}
Without inital perceptions of the time simulations 
needed to evaluate agents, expectations are that the 
experiemnt will be optimized for running a single configuration
set thoroughly.

\subsection{Documentation}
The configuration and the execution of the 
experiment will be documented, however, documntation 
of the code will be of lower priority than the remaining tasks. 



\end{document}
